A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes

### 2017-10-18 paper/aosgflp.tex
Fix G to H for GLFP CDF.  Also make changes to d to g for groups.

\subsection{GLFP model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$. Then
$$\Pr(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

@@

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{equation}
f(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
\end{equation}

@@

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population-specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

@@

\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

@@

There are two sources of variation at play: the
posterior uncertainty in the parameters, which largely depends on the
sample size, and the uncertainty in future
observations conditional on the parameters. The {\em posterior
  predictive} distribution incorporates both sources.
\begin{equation*}
  p(t_{d,new}|t) = \int p(t_{d,new}|\theta_d) p(\theta_d|t_d) \, d\theta_d
\end{equation*}

We can sample from this distribution by drawing $t_{d,new}^{(s)}$ from
$\operatorname{GLFP}(\pi_d^{(s)},
\mu_1^{(s)},\sigma_1^{(s)},\mu_{2,d}^{(s)},\sigma_{2,d}^{(s)})$, for
$s=1,...,S$, using the saved posterior draws for the model
parameters. The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
$$\frac{1}{S} \sum_{s=1}^{S} t_{d,new}^{(s)}.$$

@@

Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
relative to a new unit. We can rank the drive-models by 
$$\op{E}(L(t_{d,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{d,new}^{(s)}).$$

Posterior credible intervals for $L(t_{d,new})$ are shown in the left panel of Figure \ref{fig5}. A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel. The panel above shows credible intervals for MTTF. We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric. The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most. However, given the discrepancy of the model fit for Drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.

@@

\appendix
\section{Truncation adjusted Kaplan-Meier estimate of lifetime}
\label{sec:trunc-adj}
We first start with a nonparametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each drive-model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by

$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_d(t)}_{KM},\; t>t_{d,\text{min}}^L.$$

>>>>

\subsection{GLFP model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$. Then
$$\Pr(T \le t) = H(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

@@

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{equation}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
\end{equation}

@@

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population-specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

@@

\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

@@

There are two sources of variation at play: the
posterior uncertainty in the parameters, which largely depends on the
sample size, and the uncertainty in future
observations conditional on the parameters. The {\em posterior
  predictive} distribution incorporates both sources.
\begin{equation*}
  p(t_{g,new}|t) = \int p(t_{g,new}|\theta_g) p(\theta_g|t_g) \, d\theta_g
\end{equation*}

We can sample from this distribution by drawing $t_{g,new}^{(s)}$ from
$\operatorname{GLFP}(\pi_g^{(s)},
t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)})$, for
$s=1,...,S$, using the saved posterior draws for the model
parameters. The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$

@@

Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
relative to a new unit. We can rank the drive-models by 
$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)}).$$

Posterior credible intervals for $L(t_{g,new})$ are shown in the left panel of Figure \ref{fig5}. A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel. The panel above shows credible intervals for MTTF. We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric. The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most. However, given the discrepancy of the model fit for Drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.

@@

\appendix
\section{Truncation adjusted Kaplan-Meier estimate of lifetime}
\label{sec:trunc-adj}
We first start with a nonparametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $g$, denoted by
$\widehat{F_g(t)}_{KM}$, is conditional on survival up to
$t_{g,\text{min}}^L$, the shortest reported running time of all units
of drive-model $g$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each drive-model we select
$t_{g,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$\Pr(T>t_{g,\text{min}}^L|\theta_g)$ (the probability a hard drive has
survived up to $t_{g,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{g,\text{min}}^L|\theta_g)$. We compute the adjusted estimate by

$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_g(t)}_{KM},\; t>t_{g,\text{min}}^L.$$
